### 2021-05-27 article/manuscript.tex
Writing.

\RequirePackage[l2tabu,orthodox]{nag} % turn on warnings because of bad style
\documentclass[a4paper,11pt,bibtotoc]{scrartcl}
\usepackage{natbib}

\usepackage[utf8]{inputenc}

\usepackage[T1]{fontenc}        % Tries to use Postscript Type 1 Fonts for better rendering
\usepackage{lmodern}            % Provides the Latin Modern Font which offers more glyphs than the default Computer Modern
\usepackage{amsmath} % Provides all mathematical commands
\usepackage{amsfonts} % Provides all mathematical commands

\usepackage{hyperref}           % Provides clickable links in the PDF-document for \ref
\usepackage{grffile}            % Allow you to include images (like graphicx). Usage: \includegraphics{path/to/file}

% Allows to set units
\usepackage[ugly]{units}        % Allows you to type units with correct spacing and font style. Usage: $\unit[100]{m}$ or $\unitfrac[100]{m}{s}$

% Additional packages
\usepackage{url}                % Lets you typeset urls. Usage: \url{http://...}
\usepackage{breakurl}           % Enables linebreaks for urls
\usepackage{xspace}             % Use \xpsace in macros to automatically insert space based on context. Usage: \newcommand{\es}{ESPResSo\xspace}
\usepackage{xcolor}             % Obviously colors. Usage: \color{red} Red text
\usepackage{booktabs}           % Nice rules for tables. Usage \begin{tabular}\toprule ... \midrule ... \bottomrule

% Source code listings
\usepackage{listings}           % Source Code Listings. Usage: \begin{lstlisting}...\end{lstlisting}
\lstloadlanguages{python}       % Default highlighting set to "python"

\begin{document}

\title{GPROF-NN: A neural-network based implementation of the Goddard Profiling Algorithm}
\date{\today}

\maketitle

\section{Introduction}
\section{Methods and data}

This section describes the original GPROF algorithm as well as the two newly
developed neural-network-based versions.

\subsection{The original GPROF algorithm}

The original GPROF algorithm, here referred to simply as GPROF, is based on a
Bayesian formulation of the retrieval problem. Given an observation
$\mathbf{y}$, the result of the retrieval is defined as the posterior
distribution $P(\mathbf{x} | \mathbf{y})$ over the values $\mathbf{x}$ of a
numerical representation of the atmospheric state.

 The posterior distribution $P(\mathbf{x} | \mathbf{y})$ is calculated using
a database consisting of pairs
$(\mathbf{y}_1, \mathbf{x}_1), \ldots (\mathbf{y}_N, \mathbf{x}_N)$
of observations $\mathbf{y}_i$ and corresponding atmospheric states
$\mathbf{x}_i$. Assuming that the database describes the true a priori
distribution $P(\mathbf{x})$ sufficiently well, the expected
value w.r.t to the posterior distribution $P(\mathbf{x} | \mathbf{x})$ of the
atmospheric state $\mathbf{x}$ can be approximated using
\begin{align}\label{eq:gprof_retrieval}
  \int_{\mathbf{x} } \mathbf{x}\: dP(\mathbf{x} | \mathbf{y}) =
  \int_{\mathbf{x} } \mathbf{x}\: \frac{P(\mathbf{y} | \mathbf{x})}{P(\mathbf{y})} dP(\mathbf{x}) \approx
  \frac{\sum_i \mathbf{x} P(\mathbf{y}|\mathbf{x}_i)}{\sum_i P(\mathbf{y}|\mathbf{x})}
\end{align}
The conditional probability $P(\mathbf{y} | \mathbf{x}_i)$ of the observation
$\mathbf{y}$ given an atmospheric state from the database is modeled assuming
  unbiased  Gaussian errors in the observations:
  %
\begin{align}\label{eq:gprof_error}
  P(\mathbf{y}|\mathbf{x}_i) = \frac{1}{\sqrt{2\pi\text{det}(\mathbf{S}^{-1})}}
  \exp \left \{
  - \frac{1}{2}
  (\mathbf{y} - \mathbf{y}_i)^T \mathbf{S}^{-1}  (\mathbf{y} - \mathbf{y}_i)
  \right \}
\end{align}
for a certain covariance matrix $\mathbf{S}$ describing the errors due to noise
as well as other potential error sources that affect the comparison of a real
observation $\mathbf{y}$ and the observations in the database.
Eq.~(\ref{eq:gprof_retrieval}) corresponds to a resampling of the
states in the database with case-specific weights calculated using
Eq.~\ref{eq:gprof_error}. This can be extended to obtain specific quantiles of
the a posteriori distribution (described for example in \citet{pfreundschuh18})
or to derive probabilities of certain characteristics of the a posteriori state
such as the presence of precipitation in a given observation.

A difficulty of the formulation in Eq.~(\ref{eq:gprof_retrieval}) is that
calculating the retrieval result requires traversing a large database. To
speed up the retrieval, observations in the GPROF database are clustered
with respect to the database observations $\mathbf{y}_i$, so that for each
cluster of observations only the central observations, the corresponding
mean state and the number of observations in the cluster need to be stored.

Since the retrieval of complex atmospheric states from the limited information
content of a multi-spectral microwave observation is underconstrained, it is
desirable to incorporate ancillary information to increase its accuracy. In
GPROF the database is split into strata with respect to two-meter temperature
($T_{2m}$), total-column water vapor ($\text{TCWV}$) surface type ($\text{ST}$)
as well as an additional airlifting index for observations in mountain regions.

The basic functioning of the GPROF algorithm is illustrated in
Fig.~\ref{fig:gprof_legacy}. Panel (a) illustrates the reweighting of the
database entries according to Eq.~(\ref{eq:gprof_retrieval}) and
(\ref{eq:gprof_error}). The clustering of the observations to improve retrieval
performance and the stratification of the database to incorporate acillary data
is illustrated in panel (b) and (c), respectively.

\begin{figure}[hbpt]
  \centering
    \includegraphics[width=\textwidth]{../figures/gprof_legacy}
    \caption{The GPROF retrieval algorithm. Panel (a) illustrates the
      resampling of the observation based in the database to approximate the a
      posteriori distribution. Panel (b) exemplifies the clustering of
      observations that is performed to speed up the calculation of retrieval
      results. Panel (c) illustrates the incorporation of ancillary data in
      through application of a stratified retrieval database.}
  \label{fig:gprof_legacy}
\end{figure}

\subsection{Training and test data}

The neural-network-based implementations of the GPROF algorithm that will be
presented below use the same data for training and evaluation as the original
GPROF algorithm. This is done to ensure a fair comparison comparability of the
algorithms against each other as well as functional equivalence.

The main part of the GPROF retrieval database consists of atmospheric profiles
and corresponding observations derived from a year of combined radar-radiometer
observations from the Dual-frequency Precipitation Radar (DPR) and the GPM
Microwave Imager (GMI). These profiles are complemented with gauge-corrected
precipitation estimates derived from ground radar over the US for snow surfaces
and precipitation derived from ERA5 over sea ice.

The available data has been split into three parts: Data from day 6 to the end
of each month are used for training and data from day 4 and 5 are used to
monitor retrieval performance during training and to tune model hyperparameters.
The data from day 1 to 3 of each month is set aside and used for the final
comparison of the three algorithms presented in Sect.~\ref{sec:results}.

The various retrieval variables that are present in the database are listed
in Tab.~\ref{tab:variables}. While surface precipitation is certainly the
primary target of the GPROF algorithm, the observations are sensitive to
a range of additional quantities. Profiles are represented at 28 levels
with a resolution of $500\ \unit{m}$ between $0$ and $10\ \unit{km}$ and
$1 \ \unit{km}$ above that.

\begin{table}
  \caption{Retrieval quantities in the retrieval database}
  \label{tab:variables}
  \centering
  \begin{tabular}{l|r|r}
    Retrieval variable & Unit & Type\\
    \hline
    \hline

    Surface precipitation & $\unit{mm\ h^{-1}}$ & Scalar \\
    Convective precipitation & $\unit{mm\ h^{-1}}$ & Scalar \\
    Cloud water path & $\unit{kg\ m^{-2}}$ & Scalar \\
    Rain water path & $\unit{kg\ m^{-2}}$ & Scalar \\
    Ice water path & $\unit{kg\ m^{-2}}$ & Scalar \\
    Cloud water content & $\unit{g\ m^{-3}}$ & Profile \\
    Rain water content & $\unit{g\ m^{-3}}$ & Profile \\
    Snow water content & $\unit{g\ m^{-3}}$ & Profile \\
    Latent heat & $\unit{K \ h^{-1}}$ & Profile \\
  \end{tabular}
  \end{table}

\subsection{Probabilistic regression with neural networks}

The ill-posed nature of the retrieval of many atmospheric variables from
satellite observations requires a probabilistic approach in order to quantify
the inherent uncertainty of the retrieval results. In many retrievals these
uncertainties are handled by employing Bayesian methods. As shown in a previous
study \cite{pfreundschuh18}, instead of a direct application of Bayes theorem
the posterior distribution may also be predicted directly using neural networks
in a probabilistic regression framework.

The approach proposed in \citet{pfreundschuh18} uses a quantile regression
neural network (QRNN) to predict a sequence of quantiles of the posterior
distribution and has been show to yield equally reliable results as other
explicitly Bayesian retrieval methods that are commonly used to retrieve
atmospheric quantities. For each scalar retrieval variable $x$, the QRNN is
trained to predict a vector of values corresponding to the quantiles of the a
posteriori distribution by minimizing the sum of the quantile losses
\begin{align}
  \mathcal{L}_\tau(\hat{x}, x) &= (\tau - \mathbb{I}_{x < \hat{x}})(x - \hat{x})
\end{align}
for the selected quantile levels $\tau = \tau_1, \ldots, \tau_n$.

In addition to QRNNs, another approach for non-parametric probabilistic
regression has been considered for the GPROF-NN algorithm. This approach
consists of predicting a binned approximation of the probability density
function (PDF) of the posterior distribution. A prior application of this
approach can be found in \cite{metnet}. In lack of a better name, we will refer
to this approach as density regression neural network (DRNN). For each scalar
retrieval variable $x$, the DRNN is trained to classify the retrieval input into
a discrete sequence of bins covering the range of values of $x$. This is done
by converting values of the scalar retrieval variable $x$ into a categorical
variable and training the network using a categorical cross entropy loss:
\begin{align}
  \mathcal{L}(\hat{x}, x) &= \log(p_i) - \sum_{i \neq j}  \log(p_j)
\end{align}

Figure~\ref{fig:qrnn_drnn} illustrates these two approaches when applied to
retrieve surface precipitation along the central pixels of the GMI swath.
While the QRNN provides a direct estimate of the posterior CDF, the DRNN
provides an estimate of the corresponding PDF. Since both approaches provide
a direct estimate of the a posteriori distribution, the two approaches can
be used interchangeably.

\subsection{GPROF-NN 0D}

The GPROF-NN 0D algorithm retrieves all cloud- and precipitation-related
retrieval variables from a single observation and the ancillary data, which
makes it functionally equivalent to that algorithm.

\subsubsection{Network architecture}

\begin{figure}[hbpt]
  \centering
    \includegraphics[width=\textwidth]{../figures/gprof_nn_0d}
    \caption{Illustration of the neural network architecture used in the
      GPROF-NN 0D algorithm. The network consists of a common body and
      one head for each retrieval variable. Each block in body and head
      consists of fully-connected layer, layer norm and GELU activation function.
    }
  \label{fig:gprof_nn_0d}
\end{figure}

The network used in the GPROF-NN 0D algorithm is a feed-forward multi-layer
perceptron a  separate head for each of the different retrieval variables.
Profile variables share a single head and the expansion of the profile levels
is performed in the last layer of the network. The basic building block of
the network consists of a fully-connected layer followed by layer normalization
and a GELU activation function. The body of the network consists of
$N_b$ of those blocks and each head of $N_h$. Note that the case $N_b = 0$ is
equivalent to predicting each retrieval target with a separate neural network.
A shared body in the neural network allows computations to be shared between
different retrieval variables and therefore makes inference and training more
efficient.

\subsubsection{Training}

The GPROF-NN 0D algorithm is trained to predict all retrieval variable in
parallel by minimizing the sum of all losses. Each network is trained for
for 70 epochs with the Adam optimizer and a cosine annealing learning rate
schedule. Warm restarts are performed after 10, 20 and 40 epochs. The initial
learning rates for the first and second part of the training set to $5\cdot 10^{-4}$
and reduced to $10^{-4}$ for its last part.

learning rate is

\subsubsection{Validation-set performance}

\subsection{GPROF-NN 2D}

\subsubsection{Architecture}

\subsubsection{Validation-set performance}

\section{Results}

\bibliographystyle{alpha}
\bibliography{research}

\end{document}