### 2021-05-28 article/manuscript.tex

\title{GPROF-NN: A neural-network based implementation of the Goddard Profiling Algorithm}

\documentclass[12pt]{article}

\usepackage{amsmath}
\begin{document}
\maketitle

\section{Introduction}
\section{Methods}

This section describes the original GPROF algorithm as well as the two newly
developed neural-network-based versions.

\subsection{The original GPROF algorithm}

The original GPROF algorithm is based on a Bayesian formulation of the retrieval
problem. Given an observation $\mathbf{y}$, the result of the retrieval is
defined as the posterior distribution $P(\mathbf{x} | \mathbf{y})$ of the
atmospheric state $\mathbf{x}$.

To calculate the posterior $P(\mathbf{x} | \mathbf{y})$ the algorithm uses a
database consisting of pairs
$(\mathbf{y}_1, \mathbf{x}_1), \ldots (\mathbf{y}_N, \mathbf{x}_N)$
of observations $\mathbf{y}_i$ and corresponding states
$\mathbf{x}_i$. Assuming that the database describes the true a priori
distribution of atmospheric states $\mathbf{x}$ sufficiently well, the expected
value w.r.t to the posterior distribution $P(\mathbf{x} | \mathbf{x})$ of the
atmospheric state $\mathbf{x}$ can be approximated using
\begin{align}
  \int_{\mathbf{x} } f(\mathbf{x})\: d\mathbf{x} \approx
  \frac{\sum_i f(\mathbf{x}) P(\mathbf{y}|\mathbf{x}_i)}{\sum_i P(\mathbf{y}|\mathbf{x})}
\end{align}
The conditional probability of the observation given an atmospheric state
from the database, $P(\mathbf{y}|\mathbf{x}_i)$, is approximated using a
Gaussian error model
\begin{align}
  P(\mathbf{y}|\mathbf{x}_i) = \frac{1}{\sqrt(2\pi\text{det}(\mathbf{S}^{-1}))}
  \exp \left \{
  - \frac{1}{2}
  (\mathbf{y} - \mathbf{y}_i})^T \mathbf{S}^{-1}  (\mathbf{y} - \mathbf{y}_i)}{2})
  \right }
\end{align}

This approach can be extended to also estimate other statistics of the a
posteriori distribution such as its quantiles.

\end{document}