### 2021-12-11 article/manuscript.tex
Manuscript writing.

\RequirePackage[l2tabu,orthodox]{nag} % turn on warnings because of bad style
\documentclass[a4paper,11pt,bibtotoc]{scrartcl}
\usepackage{natbib}

\usepackage[utf8]{inputenc}

\usepackage[T1]{fontenc}        % Tries to use Postscript Type 1 Fonts for better rendering
\usepackage{lmodern}            % Provides the Latin Modern Font which offers more glyphs than the default Computer Modern
\usepackage{amsmath} % Provides all mathematical commands
\usepackage{amsfonts} % Provides all mathematical commands

\usepackage{hyperref}           % Provides clickable links in the PDF-document for \ref
\usepackage{grffile}            % Allow you to include images (like graphicx). Usage: \includegraphics{path/to/file}
\usepackage{subcaption}

% Allows to set units
\usepackage[ugly]{units}        % Allows you to type units with correct spacing and font style. Usage: $\unit[100]{m}$ or $\unitfrac[100]{m}{s}$

% Additional packages
\usepackage{url}                % Lets you typeset urls. Usage: \url{http://...}
\usepackage{breakurl}           % Enables linebreaks for urls
\usepackage{xspace}             % Use \xpsace in macros to automatically insert space based on context. Usage: \newcommand{\es}{ESPResSo\xspace}
\usepackage{xcolor}             % Obviously colors. Usage: \color{red} Red text
\usepackage{booktabs}           % Nice rules for tables. Usage \begin{tabular}\toprule ... \midrule ... \bottomrule

% Source code listings
\usepackage{listings}           % Source Code Listings. Usage: \begin{lstlisting}...\end{lstlisting}
\lstloadlanguages{python}       % Default highlighting set to "python"

\renewcommand{\familydefault}{\sfdefault}
\newcommand\todo[1]{\textcolor{red}{\textbf{TODO :: }#1}}
\newcommand\note[1]{\textcolor{green}{\textbf{NOTE :: }#1}}

\begin{document}

\title{GPROF-NN: A neural-network based implementation of the Goddard Profiling Algorithm}
\date{\today}

\maketitle

\section{Introduction}

The Goddard Profiling Algorithm (GPROF, \citet{kummerow15}) is the operational
precipitation retrieval algorithm for the passive microwave (PMW) observations
from the constellation of satellites of the Global Precipitation Measurement
(GPM, \citet{hou14}) mission, whose objective is to provide global measurements
of precipitation at a temporal resolution of 3 hours. The precipitation that is
retrieved using GPROF is used directly by meteorologists and climate scientists
and serves as input for GPM level 3 retrieval products. The algorithm thus
constitutes an essential component of the system of global observations that
enables monitoring of the hydrological cycle for the benefit of science and
society.

The development of GPROF was originally motivated by the Tropical Rainfall
Measurement Mission (TRMM, \citet{simpson96}), the precursor of the GPM mission,
and thus dates back almost 30 years \citep{kummerow94_a, kummerow94_b}. Due to
the conceptual and computational complexity of simulating PMW observations of
clouds and precipitation, the algorithm was and remains based on a retrieval
database consisting of observations and corresponding profiles of hydrometeors
and precipitation rates. Nonetheless, the algorithm has undergone several
updates since its conception: Methodologically, the most fundamental
modification was the introduction of the Bayesian retrieval scheme in
\citet{kummerow96}, which is used in the algorithm until today. Following this,
algorithm updates were mostly focused on improving the retrieval database and
the incorporation of ancillary data into the retrieval. While the first version
of GPROF still used hand crafted hydrometeor profiles to generate the retrieval
database, these were soon replaced by profiles from a meso-scale weather model
\citep{kummerow96}. An important improvement was the replacement of the
model-derived database by an observationally-generated database for the GPROF
2010 algorithm \citep{kummerow11, kummerow15}, which helped reduce errors caused
by misrepresentation of atmospheric states in the database. The 2014 version of
GPROF \citep{kummerow15} introduced the first fully-parametric version of the
algorithm, which was designed to be applicable to all sensors of the GPM
constellation. This version of GPROF became the operational PMW
precipitation retrieval of the GPM mission.

This study focuses on the computational method that is used to produce the
retrieval results from the retrieval database used by GPROF. Since its
introduction in \citet{kummerow96}, the currently used Bayesian method has not
received much consideration, mainly because the database and the incorporation
of ancillary data were deemed to be more relevant for improving the accuracy of
the retrieval. However, two disadvantages of the current retrieval method have
become apparent with the introduction of the much larger,
observationally-generated retrieval databases into the algorithm
\citep{elsaesser15}: Firstly, the retrieval database must be compressed into
self-similar clusters to make calculation of the retrieval results sufficiently
fast. This lossy compression may limit the extent to which the current algorithm
can benefit from the size and representativeness of observationally generated
retrieval databases. Secondly, the current retrieval method requires artificial
inflation of the assumed database uncertainties to produce accurate results.
Since there is no principled way to calculate these uncertainties, they need to
be tuned heuristically for each sensor.

With the aim of overcoming these limitations, this study investigates the
benefits of using the GPROF retrieval database to train a neural network to
retrieve precipitation and hydrometeor profiles. Since the retrieval database
has grown to a size of several hundred million entries it is perfectly suited
for the application of deep neural networks, which scale very well to large
amounts of data and are capable of learning highly complex relationships from
them. In addition to this, a neural network based implementation has the
advantage of allowing the integration of spatial information into the retrieval,
which is not easily feasible with the current method. This study thus aims to
answer the following two questions:
\begin{enumerate}
\item Can a deep learning based retrieval method with identical inputs improve
  the accuracy of retrieved surface precipitation and vertical hydrometeor
  profiles?
  \item Can the incorporation of spatial information into the retrieval help to
    improve the retrievals?
\end{enumerate}

To answer these questions, this study presents two novel, neural network based
implementations of the GPROF algorithm:
\begin{description}
\item[GPROF-NN 1D:] This algorithm uses a fully-connected neural network to
  retrieve single column hydrometeors and rain rates based on the observed
  brightness temperature vector. Its inputs and outputs are thus identical to
  those of the standard GPROF algorithm.
\item[GPROF-NN 3D:] This algorithm extends the GPROF-NN 1D algorithm by
  incorporating spatial information into the retrieval. It produces the same
  output as GPROF and GPROF-NN 1D but processes all observations simultaneously,
  thus allowing the algorithm to combine information from pixels across the
  swath.
\end{description}
The retrieval performance of the two GPROF-NN implementations is assessed on a
held-out part of the retrieval database and compared to that of the upcoming
version of GPROF. The performance is assessed for two sensors of the GPM
constellation: The GPM Microwave Imager (GMI) and the Microwave Humidity Sounder
(MHS, \citet{bonsignori07}). In addition to this, a case study is presented that
compares the retrieved surface precipitation from overpasses of the two sensors
over Hurricane Harvey.

\section{Data and methods}

This section introduces the retrieval database that is used by the GPROF
algorithm and serves as training data for the GPROF-NN algorithms. This
is followed by a description of the current implementation of GPROF as well
as the implementation of the two GPROF-NN retrievals.

\subsection{The retrieval database}

The GPROF retrieval database consists of the retrieval inputs, i.e. PMW observations
and corresponding ancillary data, paired with corresponding retrieval outputs,
i.e. the values of the retrieval quantities. GPROF retrieves surface
precipitation, profiles and path integrals of other hydrometeors as well as
latent heating profiles. A listing of all retrieval targets and corresponding
units is provided in Tab.~\ref{tab:variables}.

\begin{table}
  \caption{Retrieval quantities in the retrieval database}
  \label{tab:variables}
  \centering
  \begin{tabular}{l|rr}
    Retrieval variable & Unit & Type\\
    \hline
    Surface precipitation & $\unit{mm\ h^{-1}}$ & Scalar \\
    Convective precipitation & $\unit{mm\ h^{-1}}$ & Scalar \\
    Cloud water path & $\unit{kg\ m^{-2}}$ & Scalar \\
    Rain water path & $\unit{kg\ m^{-2}}$ & Scalar \\
    Ice water path & $\unit{kg\ m^{-2}}$ & Scalar \\
    Cloud water content & $\unit{g\ m^{-3}}$ & Profile \\
    Rain water content & $\unit{g\ m^{-3}}$ & Profile \\
    Snow water content & $\unit{g\ m^{-3}}$ & Profile \\
    Latent heating & $\unit{K \ h^{-1}}$ & Profile \\
  \end{tabular}
\end{table}

Values of the retrieval targets for the database of the upcoming version of
GPROF are derived from one year (October 2018 to September 2019) of
hydrometeor-profile retrievals from the GPM combined product \citep{grecu16}.
This data is complemented with surface precipitation from the currently
operational MIRS retrieval by adding light precipitation in areas where no echo
is detected by the GPM Dual-Frequency Precipitation Radar. Observations over
sea-ice and snow-covered surfaces are handled separately. For sea-ice,
precipitation is derived from the ERA5 reanalysis \citep{hersbach20}. For
snow-covered surfaces, precipitation is derived from four years of co-locations
with the gauge-corrected radar observations from the Multi-Radar Multi-Sensor
(MRMS) product.

The retrieval targets are then combined with observations from each of the
sensors of the GPM constellation: For GMI, this is done by simply remapping and
averaging the retrieval outputs to the footprint of GMIs $23\ \unit{GHz}$
channel. Since co-locations with other sensors of the constellation are rare or
may not by available at all, observations for these sensors are simulated and
the footprint averaging is adapted to the expected sensor resolution.

Furthermore, ancillary data derive from global circulation models is added to
the database. It comprises two meter temperature ($T_{\text{2m}}$), total column
water vapor ($\text{TCWV}$), a surface class as well as an air lifting index
(ALI) that encodes information on atmospheric convergence in mountainous areas.
Retrieval databases are produced in two configurations for each sensor of the
GPM constellation: One for near real time applications, which uses forecasts
from the Japan Meteorological Agency (JMA) Global Analysis (GANAL) to derive
ancillary data, and one for climatological applications, which uses the ERA5
reanalysis \citep{hersbach20} to derive the ancillary data. Since this is more
of an operational detail, this study only uses databases based on ERA5.

\subsection{The GPROF algorithm}
\label{sec:gprof}

The current implementation of the GPROF retrieval consists of three components:
(1) The Bayesian method that is used to calculate the retrieval results from
retrieval inputs, (2) the compression required to handle very large databases
and (3) the incorporation of ancillary data into the retrieval. An illustration
of these three components is provided in Fig.~\ref{fig:gprof_legacy}.

\begin{figure}[hbpt!]
  \centering \includegraphics[width=\textwidth]{../figures/gprof_legacy}
  \caption{The GPROF retrieval algorithm. Panel (a) illustrates the resampling
    of database samples based on the Bayesian scheme that is the foundation of
    the GPROF retrieval. The colored, filled curve on the right side of the plot
    depicts the approximate PDF derived from the sample weights. Panel (b) shows
    the clustering of observations that is performed to speed up the calculation
    of retrieval results. Panel (c) illustrates the incorporation of ancillary
    data through the application of a stratified retrieval database.}
  \label{fig:gprof_legacy}
\end{figure}

The Bayesian scheme that is used in GPROF calculates the retrieval result as the
posterior distribution $p(\mathbf{x} | \mathbf{y})$ of the observed atmospheric
state $\mathbf{x}$ conditioned on the retrieval input $\mathbf{y}$, here a
vector of PMW observations. The database that is used for that consists of pairs
$(\mathbf{y}_1, \mathbf{x}_1), \ldots (\mathbf{y}_N, \mathbf{x}_N)$ of
observations $\mathbf{y}_i$ and states $\mathbf{x}_i$. Assuming the database is
distributed according to the the true a priori $p(\mathbf{x})$, i.e. the
climatological distribution of the atmospheric state $\mathbf{x}$, the expected
value of $\mathbf{x}$ w.r.t to the posterior distribution $p(\mathbf{x} |
\mathbf{x})$ can be approximated using
\begin{align}\label{eq:gprof_retrieval}
  \int_{\mathbf{x} } \mathbf{x} p(\mathbf{x} | \mathbf{y})\: d\mathbf{x} =
  \int_{\mathbf{x} } \mathbf{x}\: \frac{p(\mathbf{y} |
    \mathbf{x})p(\mathbf{x})}{p(\mathbf{y})}\: d\mathbf{x} \approx \frac{\sum_i
    \mathbf{x} P(\mathbf{y}|\mathbf{x}_i)}{\sum_i P(\mathbf{y}|\mathbf{x})}
\end{align}
The conditional probability $P(\mathbf{y} | \mathbf{x}_i)$ of the observation
$\mathbf{y}$ given an atmospheric state from the database is modeled assuming
unbiased Gaussian errors in the observations 
  %
\begin{align}\label{eq:gprof_error}
  P(\mathbf{y}|\mathbf{x}_i) = \frac{1}{\sqrt{2\pi\text{det}(\mathbf{S}^{-1})}}
  \exp \left \{ - \frac{1}{2} (\mathbf{y} - \mathbf{y}_i)^T \mathbf{S}^{-1}
  (\mathbf{y} - \mathbf{y}_i) \right \}
\end{align}
using a diagonal covariance matrix $\mathbf{S}$ that descibes the errors due to
sensor noise as well as other causes of deviations of real observation
$\mathbf{y}$ from the observations in the database.
Eq.~\ref{eq:gprof_retrieval} corresponds to a resampling of the states in the
database with case-specific weights calculated using Eq.~\ref{eq:gprof_error}.
This can be extended to approximate the probability density function of the
posterior distribution or to derive probabilities of certain characteristics of
the a posteriori state such as the presence of precipitation in a given
observation.

With the current size of the retrieval database, including all samples in the
sum in Eq.~\ref{eq:gprof_retrieval} would make the retrieval prohibitively slow.
To speed up the calculation observations are grouped into self-similar clusters
and only mean observation and state as well as the number of observations is
stored in the database. The clustering is accounted for in
Eq.~\ref{eq:gprof_retrieval} by multiplying the weight
$p(\mathbf{y_i}|\mathbf{x_i})$ with the number of profiles in the cluster.

Finally, in order to constrain the retrieval further than what the PMW
observations can do alone, the sum in Eq.~\ref{eq:gprof_retrieval} is limited to
profiles from similar meteorological contexts as identified by the ancillary
data. To this end the database is stratifed with respect to $T_\text{2m},
\text{TCWV}$, surface type and air lifting index. In addition to further
constraining the retrieval, this has the benefit of further speeding up the
retrieval.

\subsection{The GPROF-NN algorithms}

The principal objective that guided the design of the GPROF-NN algorithms was to
develop a neural network retrieval that provides the exact same output as GPROF
so that it can potentially replace the current implementation in a future
update. GPROF produces several outputs that are probabilistic: A probability of
precipitation as well as the mode and terciles of the posterior distribution of
precipitation. An implementation based on standard regression neural networks
would provide no principled way to produce these probabilistic outputs due to
incompatibility of (deterministic) regression with the Bayesian formulation used
by GPROF. To overcome this limitation, the implementation of the GPROF-NN
retrievals is based on quantile regression neural networks (QRNNs). As has been
shown in \citep{pfreundschuh18}, when trained on a dataset that is distributed
according to the a priori distribution of a Bayesian retrieval, QRNNs learn to
the predict quantiles of the Bayesian posterior distribution. They thus provide
a straight forward way to reconcile neural network retrievals with the Bayesian
framework that is applied by GPROF.

This approach can be generalized to predict a sequence of quantiles, from the
cumulative distribution function (CDF) of the a posteriori distribution can be
reconstructed. Since the distribution of a scalar variable is fully described by
its CDF, the predicted CDF can be used to derive any relevant statistic of the a
posteriori distribution. For the GPROF-NN retrievals the predicted CDF is used
to derive maximum a posteriori and mean surface precipitation (the latter of
which is identical to the solution that would have been obtained with common
mean squared error regression), the terciles of the posterior distribution as
well the probability of precipitation. An illustration this principle of the
GPROF-NN retrievals is provided in Fig.~\ref{fig:gprof_nn_principle}: The
retrieval employs a neural network to predict for each input pixel a vector of
values. The elements of this vector correspond to a specific squence of
quantiles of the a posteriori distribution. These quantiles are used to
reconstruct a piece-wise linear approximation of the CDF of the a posteriori
distribution from which the retrieval outputs are derived.

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=1.0\textwidth]{../plots/gprof_nn_principle}
  \caption{The basic principle of the implementation of the GPROF-NN retrievals.
    A Bayesian solution of the retrieval is obtained by predicting, for each
    input pixel, a sequence of quantiles of the a posterior distribution that is
    used to reconstruct its CDF. The predicted CDF is then used to derive the
    retrieval results.}
  \label{fig:gprof_nn_principle}
\end{figure}

\subsubsection{Training objectives}
\label{sec:objectives}

A neural network can learn to predict a quantiles $\hat{x}_\tau$ of
a given conditional distribution by training it to minimize the quantile (or pinball) loss
$\mathcal{L_\tau}$ function corresponding to the quantile
fraction $\tau$ \citep{koenker01}:
\begin{align}\label{eq:quantile_loss}
  \mathcal{L}_\tau(\hat{x}_\tau, x) &= (\tau - \mathbb{I}_{x < \hat{x}_\tau})(x
  - \hat{x}_\tau),
\end{align}
where $\hat{x}_\tau$ is the predicted quantile, $x$ is a sample of the
conditional distribution from the training data and $\mathbb{I}$ is the
indicator function taking the value 1 when $x < \hat{X}$ and 0 otherwise.

This principle can be extended to a sequence of quantiles corresponding to
quantile fractions $\tau_1, \ldots, \tau_N$ by minimizing the mean of the losses
functions corresponding to each quantile fraction:
\begin{align}\label{eq:loss_function}
  \mathcal{L}_{\tau_1,\ldots, \tau_N}(\hat{\mathbf{x}}, x) &= \frac{1}{N} \sum_i
  \mathcal{L}_{\tau_i}(\hat{x}_i, x)
\end{align}
where $\hat{x}_i$ is the $i$th component of the vector of predicted quantiles
$\hat{\mathbf{x}}$.

The GPROF-NN retrievals use the loss function Eq.~\ref{eq:loss_function} with
128 equally spaced quantiles ranging from $\tau_1 = 0.001$ to $\tau_{128} =
0.999$ for all scalar retrieval variables.

Hydrometeor profiles, however, are handled differently: To reduce the number of
network outputs, the posterior mean of hydrometeor profiles is predicted
directly using mean squared error regression. Since the output of GPROF contains
only the posterior mean of the hydrometeor concentrations, it was deemed
unnecessary to predict their full posterior distribution at each level using
quantile regression.

\subsubsection{GPROF-NN 1D}

The GPROF-NN 1D algorithm was designed to be functionally identical to the
current GPROF algorithm. Just as GPROF, it uses PMW observations from a single
pixel together with corresponding ancillary data to retrieve corresponding
precipitation and hydrometeor profiles.

\begin{figure}[hbpt]
  \centering
    \includegraphics[width=\textwidth]{../plots/gprof_nn_0d_arch}
    \caption{Illustration of the neural network architecture used in the
      GPROF-NN 1D algorithm. The network consists of a common body and
      one head for each retrieval variable. Each block in body and head
      consists of fully-connected layer, layer norm and GELU activation function.
    }
  \label{fig:gprof_nn_1d}
\end{figure}

The neural network architecture used for the GPROF-NN 1D retrieval is
illustrated in Fig.~\ref{fig:gprof_nn_1d}. A single network is trained to
predict all retrieval variables listed (c.f. Tab.~\ref{tab:variables}) using the
training objectives described Sec.~\ref{sec:objectives}. The network consists of
a shared body and a separate head for each retrieved variable. Body and heads
are built-up of blocks consisting of a fully-connected layer followed by
layer normalization \citep{ba16} and GELU \citep{hendrycks16} activation
functions. During development we have experiment with different numbers of blocks
in body ($N_b$) and each of the heads ($N_h$) but found only marginal impact on
the retrieval performance and settled for a configuration with $N_b = 6$ and
$H_h = 4$.

The GPROF-NN 1D network is trained by simultaneously minimizing the sum of the
losses of all retrieval variables. The training is performed over 60 epochs
using the Adam optimizer \citep{kingma14} with an initial learning rate of
$5\cdot10^{-4}$ and a cosine annealing learning rate schedule
\citep{loshchilov16}. Warm restarts are performed after 10, 20 and 30 epochs.

\subsubsection{GPROF-NN 3D}

The GPROF-NN 3D retrieval extends the GPROF and GPROF-NN 1D algorithms by
incorporating spatial information into the retrieval. To achieve this,
the GPROF-NN employs a convolutional neural network (CNN) that performs
the retrieval for all pixels in the swath simultaneously.

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=1.0\textwidth]{../plots/gprof_nn_2d_arch_compl}
  \caption{
    The neural network architecture of the GPROF-NN 3D retrieval. Illustration
    (a) displays the structure of the Xception blocks \citep{chollet17} that
    form the building blocks of the GPROF-NN 3D model. An Xception block
    consists of two depthwise separable convolutions followed by a group
    normalization layer and a GELU activation function. Illustration (b) shows
    how the Xception blocks are used in an asymmetric encoder-decoder structure
    that forms the body of the network. Output from the body is combined with
    the ancillary data to form the inputs to the separate heads that predict the
    retrieval results for each of the retrieved variables.
  }
  \label{fig:gprof_nn_3d_arch}
\end{figure}

The network architecture for the GPROF-NN 3D algorithm, illustrated in
Fig.~\ref{fig:gprof_nn_3d_arch}, consists of an asymmetric encoder-decoder
structure followed by a separate head for each retrieved variable. The stages of
the en- and decoder are built up of what we refer to here as Xception blocks
(Fig.~\ref{fig:gprof_nn_3d_arch} (a)) because they are based on the Xception
architecture introduced in \citet{chollet17}. Each block consists of two
depthwise separable convolutions with a kernel size of 3 followed by group
normalization layers with 32 groups and GELU activation functions. The first
block in each stage of the encoder additionally contains a $3\times3$
max-pooling layer with a stride of 2 following the first $3\times3$ convolution
layer. Each downsampling block in the encoder is followed by $N = 4$ standard
Xception blocks. The stages of the decoder consist of bi-linear upsampling layer
followed by a single Xception block. The network architecture was chosen in
order to keep the memory footprint of the model low enough that a moderately
deep model could still be trained on a single GPU with $24\ \unit{GB}$ of
memory.

\begin{figure}[hbpt]
  \centering
    \includegraphics[width=\textwidth]{../plots/data_augmentation}
    \caption{
      The effect of GMIs conical viewing geometry on observed features. Panel
      (a) displays geolocated observations of the $10.6\ \unit{GHz}$ channel.
      Grey squares mark equilaterals with a side length of $200\unit{km}$
      oriented along the swath. The highlighted stripe located at the swath
      center marks the region where the values of the retrieved variables are
      available. Panel (b) shows the same observations viewed as an image on a
      uniform grid. Panel (c) shows six synthetically generated training inputs
      based on two input regions marked in Panel (b). The first row shows three
      synthetic samples that simulate the effect of viewing the input in region
      A at a different position across the GMI swath. The second row shows the
      corresponding transformations for the input in region B.
    }
  \label{fig:data_augmentation}
\end{figure}

Due the special structure of the GPROF retrieval database, it was necessary to
adapt the training scheme for the GPROF-NN 3D retrieval. For the most part, the
retrieval database contains values for the retrieved variables only at the 20
central pixels of the GMI swath (illustrated by the light stripe in
Fig~\ref{fig:data_augmentation} (a)). That means that although the network
predicts the retrieval variables at all pixels in the swath, the loss can be
calculated only at the pixels where the reference data is available.
Furthermore, the viewing geometry of PMW sensors breaks the translation symmetry
that is common for digital images and which is one of the inductive biases of
CNNs \citep{goodfellow16}. The reason for this is that the distance between and
orientation of pixels varies across the swathes of the sensors. This is
illustrated in Fig.~\ref{fig:data_augmentation} for the conical viewing geometry
of GMI. When displayed on a uniform grid, one and the same shape appears
differently depending on its position across the swath. To account for this in
the training, training samples are transformed randomly to simulate the effect
of seeing a given scene at different locations across the GMI swath. These
transformations are applied randomly when a training sample is loaded thus
ensuring that the network rarely or never sees a training scene from the same
perspective. The transformations were furthermore designed to ensure that the
relative location of the pixels at which values of retrieval variables are known
varies across the full width of the swath instead of being located always at its
center. Examples of the transformed inputs are displayed in
Fig.~\ref{fig:data_augmentation} (c).

The training of the GPROF-NN 3D algorithm for other sensors than GMI is further
complicated by the fact that the retrieval database for these sensors contains
simulated observations only at the central pixels of the GMI swath (the
highlighted pixels in Fig.~\ref{fig:data_augmentation} (a)). To obtain
two-dimensional training scenes that are sufficiently wide to train a CNN, we
make use of an intermediate CNN based model to predict the simulated brightness
temperatures across the full GMI swath. The extended simulated brightness
temperatures are then remapped from the GMI viewing geometry to the viewing
geometry of the target sensor in the same way as outlined in the paragraph
above.

\section{Results}

The first part of this section presents the the evaluation GPROF and the
GPROF-NN algorithms on a held-out test data set. The test data set consists of
the observations from the retrieval database from the first three days of every
month. It should be noted that we have deliberately limited this evaluation to
data from the retrieval database in order to isolate the effect of the retrieval
algorithm from that of the database.

We conclude this section with a case study of overpasses of Hurricane Harvey.
These results are based on real observations and thus provide an indication to
what extent the performance on the retrieval database can be expected to
generalize to real observations.

\subsection{Precipitation and water columns}

As described in Tab.~\ref{tab:variables}, the scalar variables retrieved by
GPROF are surface and convective precipitation as well as the column-integrated
concentrations of cloud droplets, rain and snow, denoted as cloud water path
(CWP), rain water path (RWP) and ice water path (IWP), respectively. Scatter
plots of the retrieval results for these 5 quantities are displayed in
Fig.~\ref{fig:results_scatter_gmi} for GMI and
Fig.~\ref{fig:results_scatter_mhs} for MHS. The frequencies in all plots have
been normalized column-wise to ensure that results for high  values
remain visible.

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=\textwidth]{../plots/results_scatter_gmi}
  \caption{
    Scatter plots of scalar retrieval targets for the three retrieval algorithms
    for GMI. The row display the results for the GPROF, GPROF-NN 1D and GPROF-NN
    3D algoriths, respectively. Columns display the results for different
    retrieval targets. Frequencies in the plots have been normalized
    column-wise, i.e. per bin of the reference value.
  }
  \label{fig:results_scatter_gmi}
\end{figure}

For surface precipitation retrieved by GMI, consistent improvements in retrieval
accuracy are observed between the GPROF and GPROF-NN 1D as well as the GPROF-NN
1D and GPROF-NN 3D algorithms. The improvements are most pronounced for very
small rain rates between $10^{-2}$ and $10^{-1}\ \unit{mm\ h^{-1}}$ but are
consistent across the full range of displayed values. For convective
precipitation, the results of GPROF deviate noticeably from the diagonal.
Slightly better results are provided by the GPROF-NN 1D algorithm, while the
GPROF-NN 3D algorithm yields the best agreement with the reference data. For the
path-integrated quantities GPROF already achieves good agreement with the
reference values. Nonetheless, small but consistent improvements are observed
for the GPROF-NN 1D and GPROF-NN 3D algorithms.

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=\textwidth]{../plots/results_scatter_mhs}
  \caption{
    Same as Fig.~\ref{fig:results_scatter_gmi} but for MHS.
  }
  \label{fig:results_scatter_mhs}
\end{figure}

The results for MHS, displayed in Fig~\ref{fig:results_scatter_mhs}, paint a
similar picture. Although the overall accuracy of the retrieval is lower, the
GPROF-NN 1D retrieval consistently yields more accurate results than GPROF. Also
here the GPROF-NN 3D retrieval yields further, consistent improvements compared
to the GPROF-NN 1D retrieval.

Quantitative measures of the retrieval accuracy for all scalar retrieval targets
and the three retrieval algorithms are displayed in Tab.~\ref{tab:metrics_gmi}
for GMI and Tab.~\ref{tab:metrics_mhs} in the appendix for MHS.
The tables display bias, mean absolute error (MAE), root mean squared error
(RMSE) and the mean absolute percentage error (MAPE$_{t}$) for all test samples
events with a reference value that exceeds a quantity-specific threshold $t$.
The error metrics confirm the qualitative findings from
Fig.~\ref{fig:results_scatter_gmi} and \ref{fig:results_scatter_mhs}: The neural
network implementations outperform GPROF in terms of all considered metrics.
Moreover, the GPROF-NN 3D algorithm further improves upon the performance of
the GPROF-NN 1D algorithm. Although the errors are generally larger for MHS, the
same tendency is observed.

\begin{table}[hbpt!]

  \centering
  \caption{
    Selected error metrics and standard deviations of retrieved scalar
    variables for the three implementations of the GPROF algorithm applied to
    the GMI sensor. The standard deviations were estimated using bootstrapping
    and thus provide a measure of the uncertainty due to random sampling of the
    test data.
  }
  \label{tab:metrics_gmi}
\begin{subtable}{\textwidth} 
 \caption{Surface precipitation} 
\begin{tabular}{|c||p{3.5cm}|p{3.5cm}|p{3.5cm}|}
\hline
Metric &
\multicolumn{1}{|c}{GPROF} &
\multicolumn{1}{|c}{GPROF-NN 1D} &
\multicolumn{1}{|c|}{GPROF-NN 3D} \\
\hline\hline
Bias \hfill [$\unit{mm\ h^{-1}}$] & \hfill $ -0.0091 \pm 0.0001$ &\hfill $ -0.0019 \pm 0.0001$ &\hfill $ -0.0012 \pm 0.0001$ \\
MAE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0787 \pm 0.0001$ &\hfill $  0.0579 \pm 0.0000$ &\hfill $  0.0462 \pm 0.0001$ \\
RMSE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.4049 \pm 0.0078$ &\hfill $  0.3799 \pm 0.0078$ &\hfill $  0.3398 \pm 0.0078$ \\
MAPE$_{0.1}$ \hfill [$\unit{\%}$] & \hfill $ 64.3838 \pm 0.0681$ &\hfill $ 58.0879 \pm 0.0247$ &\hfill $ 48.1921 \pm 0.0449$ \\
\hline
\end{tabular}
\end{subtable}

\begin{subtable}{\textwidth} 
 \caption{Convective precipitation} 
\begin{tabular}{|c||p{3.5cm}|p{3.5cm}|p{3.5cm}|}
\hline
Metric &
\multicolumn{1}{|c}{GPROF} &
\multicolumn{1}{|c}{GPROF-NN 1D} &
\multicolumn{1}{|c|}{GPROF-NN 3D} \\
\hline\hline
Bias \hfill [$\unit{mm\ h^{-1}}$] & \hfill $ -0.0026 \pm 0.0001$ &\hfill $ -0.0008 \pm 0.0001$ &\hfill $ -0.0004 \pm 0.0000$ \\
MAE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0332 \pm 0.0002$ &\hfill $  0.0232 \pm 0.0000$ &\hfill $  0.0228 \pm 0.0001$ \\
RMSE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.3946 \pm 0.0094$ &\hfill $  0.3550 \pm 0.0094$ &\hfill $  0.3104 \pm 0.0094$ \\
MAPE$_{0.1}$ \hfill [$\unit{\%}$] & \hfill $ 90.7868 \pm 0.1886$ &\hfill $ 88.3793 \pm 0.1458$ &\hfill $ 82.5167 \pm 0.1474$ \\
\hline
\end{tabular}
\end{subtable}

\begin{subtable}{\textwidth} 
 \caption{Rain water path} 
\begin{tabular}{|c||p{3.5cm}|p{3.5cm}|p{3.5cm}|}
\hline
Metric &
\multicolumn{1}{|c}{GPROF} &
\multicolumn{1}{|c}{GPROF-NN 1D} &
\multicolumn{1}{|c|}{GPROF-NN 3D} \\
\hline\hline
Bias \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0009 \pm 0.0000$ &\hfill $ -0.0004 \pm 0.0000$ &\hfill $ -0.0002 \pm 0.0000$ \\
MAE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0167 \pm 0.0000$ &\hfill $  0.0118 \pm 0.0000$ &\hfill $  0.0089 \pm 0.0000$ \\
RMSE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0913 \pm 0.0032$ &\hfill $  0.0955 \pm 0.0032$ &\hfill $  0.0702 \pm 0.0032$ \\
MAPE$_{0.01}$ \hfill [$\unit{\%}$] & \hfill $ 78.5511 \pm 0.1381$ &\hfill $ 69.4666 \pm 0.0849$ &\hfill $ 59.6460 \pm 0.0470$ \\
\hline
\end{tabular}
\end{subtable}

\begin{subtable}{\textwidth} 
 \caption{Ice water path} 
\begin{tabular}{|c||p{3.5cm}|p{3.5cm}|p{3.5cm}|}
\hline
Metric &
\multicolumn{1}{|c}{GPROF} &
\multicolumn{1}{|c}{GPROF-NN 1D} &
\multicolumn{1}{|c|}{GPROF-NN 3D} \\
\hline\hline
Bias \hfill [$\unit{mm\ h^{-1}}$] & \hfill $ -0.0029 \pm 0.0000$ &\hfill $ -0.0003 \pm 0.0000$ &\hfill $  0.0000 \pm 0.0000$ \\
MAE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0194 \pm 0.0000$ &\hfill $  0.0121 \pm 0.0000$ &\hfill $  0.0087 \pm 0.0000$ \\
RMSE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.1126 \pm 0.0034$ &\hfill $  0.1118 \pm 0.0034$ &\hfill $  0.0739 \pm 0.0034$ \\
MAPE$_{0.01}$ \hfill [$\unit{\%}$] & \hfill $ 76.1349 \pm 0.1828$ &\hfill $ 66.5753 \pm 0.1338$ &\hfill $ 54.5458 \pm 0.0679$ \\
\hline
\end{tabular}
\end{subtable}

\begin{subtable}{\textwidth} 
 \caption{Cloud water path} 
\begin{tabular}{|c||p{3.5cm}|p{3.5cm}|p{3.5cm}|}
\hline
Metric &
\multicolumn{1}{|c}{GPROF} &
\multicolumn{1}{|c}{GPROF-NN 1D} &
\multicolumn{1}{|c|}{GPROF-NN 3D} \\
\hline\hline
Bias \hfill [$\unit{mm\ h^{-1}}$] & \hfill $ -0.0018 \pm 0.0000$ &\hfill $ -0.0001 \pm 0.0000$ &\hfill $  0.0006 \pm 0.0000$ \\
MAE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0258 \pm 0.0000$ &\hfill $  0.0155 \pm 0.0000$ &\hfill $  0.0116 \pm 0.0000$ \\
RMSE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0483 \pm 0.0031$ &\hfill $  0.0387 \pm 0.0031$ &\hfill $  0.0302 \pm 0.0031$ \\
MAPE$_{0.01}$ \hfill [$\unit{\%}$] & \hfill $ 48.0013 \pm 0.0309$ &\hfill $ 29.9094 \pm 0.0188$ &\hfill $ 25.0521 \pm 0.0203$ \\
\hline
\end{tabular}
\end{subtable}
\label{tab:surface_precip_metrics}
\end{table}

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=\textwidth]{../plots/results_surface_types}
  \caption{
    Bias, RMSE, MAE and MAPE$_{0.1}$ of the retrieved surface precipitation for
    w.r.t. surface type and retrieval algorithm. The results for the GMI sensor
    are displayed in the first column, results for the MHS sensor in the
    second column. Error bars mark one standard deviation around the mean
    calculated using bootstrap and thus quantify the uncertainty due to the
    random sampling of the test data.
  }
  \label{fig:results_surface_types}
\end{figure}

Since the surface type has a considerable effect on the lower-frequency
observations used in the retrieval, its impact on the retrieval of surface
precipitation is assessed in Fig.~\ref{fig:results_surface_types}. The figure
displays the Bias, RMSE, MAE and MAPE for the different surface classes that are
distinguished in the retrieval. Even when different surface types are considered
separately, the results of the surface precipitation retrieval show the same
pattern as the scatter plots in Fig.~\ref{fig:results_scatter_gmi}: The results of
the GPROF-NN 1D retrieval are more accurate than those of the GPROF algorithm
and the results of the GPROF-NN 3D algorithm are in turn better than those of
the GPROF-NN 1D algorithm. These results are mostly consistent across all
surface types and considered metrics, however for some surface types the
GPROF-NN 1D algorithms yields lower MAPE than GPROF-NN 3D algorithm.

Fig.~\ref{fig:results_spatial} displays the spatial distribution of bias, RMSE
and MAPE for the three retrievals in $5\ \unit{^{\circ}} \times
5\ \unit{^{\circ}}$ boxes. As could be expected from the previous results, the
magnitudes of the biases of GPROF are considerably larger than for the other two
algorithms. Furthermore, the GPROF exhibits consistent biases across
geographical regions such as the tropics in South America and Africa as well as
east Asia, which is less the case for the neural network algorithms. Although
spatial distribution of the RMSE mostly reflects the global distribution of
precipitation, a gradual decrease in RMSE can be observed between the results of
GPROF and GPROF-NN 1D as well as GPROF-NN 1D and GPROF-NN 3D. More consistent
patterns are visible in the MAPE: All three retrievals exhibit increased MAPE
over land surfaces, which likely reflects the decrease in information content
due to the reduced contrast in the lower frequency channels. Over Ocean, the
MAPE is generally increased in the sub-tropics and tropics compared to higher
latitudes. Although these patterns are observed in the results of all
algorithms, a clear and globally consistent decrease in MAPE can be observed
between the GPROF, GPROF-NN 1D and GPROF-NN 3D retrievals.

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=\textwidth]{../figures/results_spatial}
  \caption{
    Spatial distributions of Bias (Column 1), RMSE (Column 2) and MAPE (Column
    3) error statistics for the GPROF (Row 1), GPROF-NN 1D (Row 2), GPROF-NN 3D
    (Row 3) algorithms in $5\ \unit{^{\circ}} \times 5\ \unit{^{\circ}} $ grid
    boxes.
  }
  \label{fig:results_spatial}
\end{figure}

\subsubsection{Predicted retrieval uncertainties and probabilistic rain detection}

In addition to quantitative precipitation estimates, GPROF produces estimates of
the first and second tercile of the posterior distribution of surface
precipitation, which provide an uncertainty estimate for the retrieved mean
surface precipitation, as well as a probabilistic classification of pixels into
raining and non-raining pixels based on an estimate probability of
precipitation. Due to their probabilistic nature, similar results can be
produced using GPROF-NN algorithms. From the predicted quantiles, the terciles
can be predicted by simply interpolating them to the fractions $\frac{1}{3}$ and
$\frac{2}{3}$, respectively. The probability of precipitation is calculated by
using the predicted posterior distribution to calculate the probability of the
retrieved surface precipitation to be larger than the smallest non-zero rain
rate in the training data.

\begin{table}
  \caption{
    Calibration of the predicted terciles of the posterior distribution of
    surface precipitation on the test set for the three retrieval algorithms and
    the GMI and MHS sensors.
  }
  \label{tab:terciles}
  \begin{subtable}{\textwidth} 
    \caption{GMI} 
    \centering
  \begin{tabular}{l|rrrr}
    \multicolumn{1}{c|}{Tercile} &
    \multicolumn{1}{c}{Nominal} &
    \multicolumn{1}{c}{GPROF} &
    \multicolumn{1}{c}{GPROF-NN 1D} &
    \multicolumn{1}{c}{GPROF-NN 3d} \\
    \hline
    First & 0.333 & 0.371 & 0.339 & 0.336 \\
    Second & 0.667 & 0.71 & 0.648 & 0.656 \\
  \end{tabular}
  \end{subtable}
  \begin{subtable}{\textwidth} 
    \caption{MHS} 
    \centering
    \begin{tabular}{l|rrrr}
      \multicolumn{1}{c|}{Tercile} &
      \multicolumn{1}{c}{Nominal} &
      \multicolumn{1}{c}{GPROF} &
      \multicolumn{1}{c}{GPROF-NN 1D} &
      \multicolumn{1}{c}{GPROF-NN 3d} \\
      \hline
      First & 0.333 & 0.392 & 0.339 & 0.336 \\
      Second & 0.667 & 0.743 & 0.648 & 0.656 \\
    \end{tabular}
  \end{subtable}

\end{table}

To assess the accuracy of the uncertainty estimates from GPROF and the GPROF-NN
algorithms, Tab.~\ref{tab:terciles} lists the calibration, i.e. the frequency
with which each tercile was larger than the true surface precipitation. Although
for both sensors and algorithms the observed frequencies are close to the
nominal frequencies $\frac{1}{3}$ and $\frac{2}{3}$, the GPROF-NN algorithms
yield calibrations that are closer to the nominal values.

\begin{figure}[hbpt!]
  \centering
  \includegraphics[width=0.75\textwidth]{../plots/results_pop}
  \caption{
    Calibration (Panel (a)) and receiver-operating characteristic (ROC, Panel(b)) for
    the predicted probability of precipitation.
  }
  \label{fig:results_pop}
\end{figure}

The quality of the rainin/non-raining classification is assessed in
Fig.~\ref{fig:results_pop}, which displays the calibration of the predicted
probability and the receiver-operating characteristic (ROC) curve. Similar as
for the terciles, the predicted probabilities are fairly well calibrated for all
algorithms and sensors. However, also here the GPROF-NN algorithms seem to
yields results slightly closer to the diagonal. The results for the ROC curves
are analogous: The GPROF-NN 1D algorithm yields better precipitation detection
than GPROF and the GPROF-NN 3D retrieval in turn yields slightly better
performance than the 1D version. In terms of classification skill, worse
performance is achieved for GMI than for MHS by all algorithms.

\subsubsection{Effective resolution}

Next, we aim to assess the impact of the retrieval method on the resolution of
the retrieved precipitation fields. For this, we adopt the approach from
\citet{guilloteau17}, who have studied the effective resolution of the previous
version of GPROF for the GMI and the TRMM Microwave Imager sensors. A
1-dimensional Haar wavelet decomposition in along-track direction over all 128
pixel long sequences in the test data is performed To calculate the effective
resolution. We do not consider observations for different surface types
separately. Following, \citet{guilloteau17} we examine energy spectra as well as
correlation coefficients and Nash-Sutcliffe (NS) efficiency of the coefficients
of the wavelet decomposition for the reference and retrieved surface
precipitation. The results are displayed in Fig.~\ref{fig:resolution}.

\begin{figure}[hbpt!]
  \centering \includegraphics[width=\textwidth]{../plots/resolution}
  \caption{ Spatial variability of retrieved and reference fields. Panel (a)
    shows the average of the total energy defined as the sum of the squared
    wavelet coefficient at different length scales for the reference and
    retrieved surface precipitation fields. Panel (b) shows the correlation
    coefficient between the coefficients of the reference and the retrieved
    field. Panel (c) shows the corresponding Nash-Sutcliffe efficiency. }
  \label{fig:resolution}
\end{figure}

An obvious difference to the results from \citet{guilloteau17} is that the
energy spectra of the reference precipitation field is not monotonically
decreasing. The reason for this is that the reference precipitation field in the
retrieval database is smoothed using an averaging filter adapted to the
footprint size of the respective sensor. The GPROF-NN 3D algorithm has the
highest variability in the retrieved precipitation field, followed by the
GPROF-NN 1D algorithm and GPROF. However, for all retrievals the variability
remains lower than that of the reference field. In terms of correlation
coefficients of the wavelet coefficients at different scales (Panel (b)), the
GPROF-NN 3D algorithm exhibits the highest correlation with the coefficients of
the reference field, again followed by the GPROF-NN 1D algorithm and GPROF. The
same pattern is observed for the NS efficiency. In term of effective resolution,
defined as the smallest scale at which the NS efficiency exceeds 0.5, the
GPROF-NN 3D algorithm for the GMI sensor achieves a resolution solution of
$13.5\ \unit{km}$, which is the distance between consecutive pixels in
along-track direction and corresponds to the smallest spatial scale that can be
resolved in this analysis. The effective resolutions for the GPROF-NN 1D
algorithm is $16.2\ \unit{km}$ and for GPROF $21.4\ \unit{km}$.

\note{I will add the results for MHS for GPROF and GPROF-NN 1D here when these
  results become available.}

\subsubsection{Profile retrieval variables}

In addition to precipitation fields and path-integrated hydrometeor
concentrations, GPROF retrieves concentration profiles of rain, snow and cloud
water as well as profiles of latent heating rates. Error statistics for these
varaibles are displayed in Fig.~\ref{fig:results_profiles}. Qualitatively, the
results are similar to those observed for the scalar retrieval variables: The
neural network retrievals yield significantly lower biases than GPROF with the
GPROF-NN 3D algorithm performing slightly better than the 1D version. Although
the relative improvements are smaller, similar patterns are observed for RMSE
and MAPE. These results are mostly consistent across metrics and retrieval
variables with the exception of the MAPE of cloud water content, for which no
improvement for the GPROF-NN 3D over the GPROF-NN 1D retrieval is observed. Also
here the retrievals based on MHS yield higher errors but the performances of the
three retrievals remain qualitatively similar to what is observed for GMI.

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=\textwidth]{../figures/results_profiles}
  \caption{
    Error statistics of the retrieved profile variables. Columns show the
    errors for the different retrieved variables, whereas rows show 
    altitude averaged Bias, RMSE and MAPE, respectively.
  }
  \label{fig:results_profiles}
\end{figure}

%\subsubsection{Retrieval results beyond the posterior mean}
%
%Most quantitative retrieval results of GPROF correspond to the mean of the a
%posteriori distribution. This is the case for the surface precipictation,
%convective precipitation, all path-integrated quantities as well as the
%retrieved profiles. In its current form, GPROF produces only 5 outputs that make
%use of information from the posterior distribution beyond the posterior mean:
%The estimated probability of precipitation, the precipitation flag, the 1st and
%3rd terciles of the surface precipitation (provided as a measure of uncertainty)
%and the most likely precipitation.
%
%
%The neural-network version of GPROF predicts the full posterior distribution for
%scalar retrieval variables and storing only the posterior mean discards a large
%part of this information. Yet, storing the full posterior distribution would
%require excessive amounts of disk storage and would likely overwhelm the average
%user. As a storage-efficient alternative, we propose to include a random sample
%from the posterior distribution in the retrieval output. Although these samples
%by construction neglect spatial correlations, the results should be able to
%truthfully reproduce the distribution of precipitation when used in a
%climatological analysis.
%
%As an illustration of the benefits of randomly-sampled retrieval estimates,
%Fig.~\ref{fig:results_zonal} displays the zonal distributions of surface
%precipitation on $1^\circ$-degree bins. The reference distribution of the test
%data (filled contours) is displayed in the background and the quantiles of the
%retrieved distributions (shaded lines) are drawn on top. Note that the color
%scaling has been chosen so that if the retrieved and reference distribution were
%similar, the shaded lines would lie on top of the boundaries between the filled
%contours in the background.
%
%The first row of panels shows the distributions for the case that the posterior
%mean is used as the observed precipitation value. In this case, the distribution
%of the observed values does not match the corresponding distribution of the test
%data well. The distribution of the observed precipitation is systematically
%biased towards light precipitation. When instead of the posterior mean random
%samples are used as precipitation estimates, the agreement between the retrieved
%and the test distribution is significantly improved. Although e discrepancies
%remain for the lower quantiles of the distribution, very good agreement is found
%for the high quantiles of the distribution.
%
%  \begin{figure}[hbpt!]
%    \centering
%    \includegraphics[width=\textwidth]{../figures/zonal_distributions}
%    \caption{
%      Zonal distributions of test data and corresponding retrieved values. Colored, filled contours
%      in the background show the CDFs of the zonal distribution of precipitation in the test data.
%      Grey-shaded line display the quantiles of the corresponding retrieved distributions based on
%      the posterior mean (Row 1) and random samples from the posterior distribution (Row 2).
%    }
%    \label{fig:results_zonal}
%  \end{figure}
%
%  For another perspective on these results, Fig.\ \ref{fig:scatter_percentiles}
%  displays scatter plots of the retrieved and the true 95th and 99th percentiles
%  of the distributions of surface precipitation on a grid of size
%  $5\ \unit{^{\circ}} \times 5\ \unit{^{\circ}}$. When the posterior mean is
%  used to calculate the distribution of surface precipitation, the obtained
%  percentiles systematically underestimate the true percentiles. This systematic
%  underestimation is corrected when instead of the posterior mean random samples
%  are used to calculate the surface precipitation distributions.
%
%  \begin{figure}[hbpt!]
%    \centering
%    \includegraphics[width=\textwidth]{../figures/scatter_percentiles}
%    \caption{
%      95th (Row 1) and 99th (Row 2) percentiles of the distribution
%      of surface precipitation in $5^\circ \times 5^\circ$ boxes in the
%      test data scattered against the corresponding retrieved values. Filled
%      contours in the background show the distribution obtained by using the
%      posterior mean as retrieved value and colored contours the distribution
%      that is obtained when a random sample from the posterior distribution
%      is used as retrieval value. The columns show the results for the
%      GPROF-NN 0D (Column 1) and the GPROF-NN 3D algorithm  (Column 2).
%    }
%    \label{fig:scatter_percentiles}
%  \end{figure}

\subsection{Case study: Hurricane Harvey}

So far, all of the presented results were based on a test dataset with the same
statistics as the retrieval database. When the algorithms are applied in
operations, however, the input data may not necessarily follow the distribution
of the retrieval database. While for GMI the observations can be expected to be
consistent with those in the retrieval database, this is not necessarily the
case for other sensors for which the retrieval database contains mostly
simulated observations. While a comprehensive analysis of the retrieval
performance on real observations is outside the scope of this paper, this
section presents retrieval results from two overpasses over Hurricane Harvey to
provide an indication as to whether the performance characteristics of the
retrieval algorithm can be expected to carry over to real observations.

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=\textwidth]{../plots/surface_precip_harvey_gmi_ir}
  \caption{ Retrieved surface precipitation from the GPM combined product
    (Panels (a), (c)), the GPROF algorithm (Panels (b), (d)), and the neural
    network versions GPROF-NN 1D (Panels (e), (g)) and GPROF-NN 3D (Panels (f),
    (h)) in and around Hurricane Harvey on 2017-08-25 at 11:50:00 UTC. Panels
    (a), (b), (e), (f) display the precipitation on a logarithmic scale over an
    area spanning $20\ 000\ \unit{km}$ in zonal and meridional directions.
    Panels (c), (d), (g), (f) display the surface precipitation on a linear
    color scale in an enlarged area spanning $5\ 000\ \unit{km}$ around the
    Hurricane. The background show $10.3\ \unit{\mu m}$ brightness temperatures
    from the GOES 16 observations closest to the overpass.
  }
  \label{fig:harvey_gmi}
\end{figure}

First, we consider an overpass of the GPM Core Observatory over Hurricane Harvey
that occurred on August 25, 2017 at 11:50 UTC. Retrieved surface precipitation
is displayed in Fig.~\ref{fig:harvey_gmi}. In addition to the retrieved
surface precipitation from GPROF and the GPROF-NN algorithms, the figure also
displays the surface precipitation from the GPM combined product averaged to the
resolution of the $18.7\ \unit{GHz}$ channel. When considered at log-scale,
obvious structural differences in the retrieved precipitation are visible: The
GPROF retrieval produces large swaths of low precipitation that cover most of
the scene, which are not present in the combined retrieval. These features are
significantly reduced in the results of GPROF-NN 1D algorithm and even more so
in the results of the GPROF-NN 3D retrieval. When considered at linear scale,
all three retrievals capture the global structure of the precipitation but
tend to overestimate its magnitude compared to the combined retrieval. This
effect is strongest for GPROF but is also apparent for the neural network
based retrievals. Nonetheless, both neural-network based retrievals better
reproduce the fine-scale structure that is visible in the results from
the combined retrieval.

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=\textwidth]{../plots/surface_precip_harvey_mhs}
  \caption{Surface precipitation retrieved from ground radar, (Panels (a), (c)),
    and MHS PMW observations using the GPROF algorithm (Panels (b), (d)), and
    the neural network versions GPROF-NN 1D (Panels (e), (g)) and GPROF-NN 3D
    (Panels (f), (h)) in Hurricane Harvey on 2017-08-25 at 13:58:00
    UTC. Panels (a), (b), (e), (f) display the precipitation on a logarithmic
    scale over an area spanning $20\ 000\ \unit{km}$ in zonal and meridional
    directions. Panels (c), (d), (g), (f) display the surface precipitation on a
    linear color scale in an enlarged area spanning $5\ 000\ \unit{km}$ around
    the Hurricane. In the background the closest true-color composite from GOES
    16 is shown. }
  \label{fig:harvey_mhs}
\end{figure}

Finally, we also present results from an overpass of the MHS sensor on board the
NOAA-18 satellite over the same storm at 13:58 UTC in
Fig.~\ref{fig:harvey_mhs}. The figure also presents radar derived
precipitation rates from NOAA's MRMS product suite as reference data. Similar as
for GMI, the GPROF algorithms predicts low precipitation rates across large
parts of the scene and even in areas that are cloud free. Albeit to a lesser
degree, this seems to be an issue also for the GPROF-NN 1D algorithm. The
GPROF-NN 3D algorithm is less affected by this issue and also yields better
agreement with the MRMS precipitation rates over land. Regarding the high rain
rates at the center of the Hurricane, the overall structure again agrees well
with the radar measurements. GPROF yields the highest rain rates, followed by
the GPROF-NN 3D and 1D algorithms. Considering the distance of the radar
observations from the coast, however, it is unclear how reliable the ground
radar measurements are.

\section{Discussion}

This study presented two novel, neural network based implementations of the
GPROF retrieval algorithm and evaluated their performance for the GMI and MHS
sensors against the current implementation.

\subsection{Retrieval performance}

The presented results show that retrieval accuracy across a wide range of
metrics (Tab.~\ref{tab:metrics_gmi}, Tab.~\ref{tab:metrics_mhs}) is consistently
improved by  $10$ to $25\ \unit{\%}$ across most retrieved quantities and
considered error metrics simply by replacing the current retrieval method with a
fully-connected neural network that uses the same input as the current GPROF
algorithm. At the same time the effective resolution of the retrieval increases
by about $25\ \unit{\%}$ from $21\ \unit{km}$ to $16\ \unit{km}$
(Fig.~\ref{fig:resolution}). Although both GPROF and the GPROF-NN 1D algorithm
are based on the same retrieval database and use the information as retrieval
input, the neural network based retrieval provides more accurate results. We
suspect that this may be cause by two aspects of the current implementation of
GPROF: Firstly, the assumption of state independent, uncorrelated Gaussian
errors that affect the observations and, secondly, the clustering of
observations to compress the retrieval database. It was found by
\citet{elsaesser15} that even small errors in the assumed uncertainties can
cause significant deviations at the pixel level. From the uncertainties that
affect the observations, only the sensor noise conforms to the Gaussian error
model employed by the retrieval method. For sensors other than GMI, the
observations are additionally affected by the modeling error in the simulated
observations, which is difficult to estimate and unlikely to be well described
by the Gaussian error model. In practice the uncertainties are additionally
inflated to account for the sparsity of the retrieval database and the effects
of the clustering. We therefore hypothesize that the ability of the GPROF-NN 1D
retrieval to learn to account for these uncertainties from the data directly
allows it to perform better than the current retrieval scheme even when both
retrievals limited to the same amount of information.

By further extending the retrieval to incorporate structural information, its
accuracy can be further improved by another $10$ to $25\ \unit{\%}$
compared to the baseline retrieval at the same time as the effective resolution
in along track direction is decreased to its lower limit of $13.5\ \unit{km}$.
The use of structural information for precipitation retrievals is common
practice in algorithms based on infrared observations \citep{sorooshian00,
  hong04} and the potential benefits of CNN based retrieval have been shown in
\citet{sadeghi19}. Our results indicate that also retrievals based on PMW
observations can benefit substantially from the incorporation of structural
information into the retrieval.

\subsection{Limitations}

It is important to consider the limitations of the results presented in this
study. We have deliberately limited the evaluation of retrieval accuracy to test
data with the same statistical properties as the training data. This was done to
isolate the effect of the retrieval method from potential aliasing effects that
would be introduced by the use of external validation data. The performance
improvements presented here must therefore be understood as upper limits on the
improvements that can be realized in operational use. Since the GMI retrieval
can be trained using real observations, the tightness of this bound depends only
on the representativeness of the retrieval database and thus the accuracy of the
sources used to generate it. For other sensors, however, the retrieval algorithm
is trained using simulated observations. Before these simulated observations can
be used to train the GPROF-NN 3D algorithm, they have to be first extended to
the full GMI swath using a dedicated simulator network and then remapped to the
viewing geometry of the sensor. Since, at least for MHS, the distribution of
simulated observations were found to not match the real observations very well,
an additional quantile-matching correction is applied to all observations. For
those sensors it is thus not at all clear to what extent the potential
improvements carry over to operational application of the retrieval.

\subsection{Operational considerations}

This study constitutes a first step towards the development of a neural network
based implementation of GPROF that may potentially replace the current
implementation. However, some challenges remain in the transition of the
GPROF-NN algorithms from experimental to operational stage. As described above,
it is not clear to what extent the benefits of the neural network based
implementation can be realized in operational use. Furthermore, there are other
aspects that affect the suitability of a retrieval for operational use. One of
them is certainly the computational effort required to perform the retrieval.
For the current neural network implementations of GPROF this is not an issue as
they in fact run faster than the current implementation even when run on a
single CPU core (c.f. Fig.~\ref{fig:timings} in the appendix). Another issue,
however, is the detection of anomalous inputs. The current implementation of
GPROF has the advantage that anomalous inputs can potentially be detected
through the missing of suitable matches from the retrieval database. The neural
network implementation lacks this capability and will produce precipitation
rates for all inputs that are not otherwise flagged by quality control. That
being said, it is not clear either how effective the current GPROF
implementation is at detecting small shifts in the retrieval input.

\subsection{Conclusions}

The results presented in this study demonstrate the potential of the two new
neural network based implementations of GPROF to improve the accuracy as well as
the effective resolution of retrievals of precipitation and hydrometeor profiles
from the PMW observations of the satellites of the GPM constellation. In
addition to the benefits of simply replacing the current retrieval scheme with a
neural network, we found similar, additional improvements by incorporating
spatial information into the retrieval through the use of a CNN based
implementation.

Nonetheless, the most relevant question remains to be answered: How well can the
retrievals generalize to real observations and independent validation data? This
will depend on the accuracy of the retrieval database as well as the realism of
the simulations that are used to produce the input data for sensors other than
GMI. While we aim to answer this question in a follow-up study, the results
presented here have value also in their own right. Should the improvements
observed here not carry over to operational use, it would point to serious
issues with the retrieval database.

Should, on the contrary, these improvements carry over to operational use these
novel GPROF implementations would constitute a simple and cost efficient way to
improve global measurements of precipitation. With this the algorithms would
constitute a small step towards improving our understanding of the global
hydrological cycle and its changes in a warming climate.

%The GPROF-NN algorithms and their performance assessment that were presented in
%this study indicate that the retrieval accuracy of the PMW precipitation retrievals
%of the GPM constellation can be improved significantly by replacing the current
%algorithm with a neural network based implementation. Although the presented
%results should be interpreted as an upper bound on the true retrieval
%performance, our results remain encouraging: Replacing the retrieval method has
%the potential to be a simple and cheap way to improve global precipitation
%measurements.
%
%The next step towards such an upgrade will be to extend the implementation of
%the GPROF-NN 0D and GPROF-NN 3D algorithms to the other sensors of the GPM
%constellation and run them operationally along side GPROF. The three methods
%will then be validated and compared against gauge corrected ground based radar
%measurements in a follow up study. This is to ensure that the good performance of
%the neural network based retrieval does indeed carries over to the operational
%processing.
%
%In addition to potential improvements in retrieval accuracy, we presented two
%novel applications enabled by neural-network based retrievals:
%
%\begin{enumerate}
%\item The inclusion of random samples from the posterior distribution in the
%  retrieval output is proposed as an efficient way of representing retrieval
%  uncertainties that helps to improve the representation of extreme events in
%  the retrieval results.
%\item Our results show that incorporating structural information into the
%  retrieval yields larger performance benefits than incorporating ancillary
%  information. This allows simplifying the retrieval pipeline to the point that
%  only observations are required as input, which opens up the possibility
%  letting the user run the retrieval on his own computer, i.e. to perform the
%  inference 'on the edge'. This may be beneficial for retrievals with very high
%  resolution, i.e. whose storage requirement are much larger than those of the
%  L1C data.
%\end{enumerate}
%
%Building on the approach proposed in \citet{pfreundschuh18}, this work presents

%pipelines built on Bayesian retrieval algorithms. We find that the machine
%learning based retrieval, is easiert to implement, runs faster and performs
%better than the hand-crafted retrieval algorithm. Although it remains to be seen
%to which extent this will improve retrievals with respect to independent
%validation data, it seems reasonable to expect at least incremental improvements
%in global precipitation measurements as well as interesting opportunities for
%the novel retrieval applications.

%\subsection{Retrieval performance}
%
%Our results show a consistent improvement of the GPROF-NN retrievals over the
%original GPROF algorithm. This is an encouraging result considering that all
%three retrievals are based on exactly the same retrieval database, indicating
%that the GPROF retrievals can be improved simply by switching to a different
%retrieval method.
%
%Nonetheless, this raises the question why the neural network based retrievals
%perform better than the original method? To investigate this, we have calculated
%the gradients of the retrieved surface precipitation with respect to the
%brightness temperatures for the GPROF and GPROF-NN 0D retrievals
%(Fig.~\ref{fig:sensitivity}). Comparison of the distributions of the gradients
%reveals that the GPROF-NN retrievals exhibit larger gradients with more
%variability than the GPROF retrieval. This shows that GPROF is less
%sensitive to the input observations than the neural network versions.
%
%Based on the formulation of GPROF (c.f. ~\ref{sec:gprof}), the retrieval may be
%viewed as nearest-neighbor regression with a Gaussian kernel and a diagonal
%covariance matrix. Although the theoretical formulation of the retrieval assumes
%that the elements of the covariance matrix to correspond to the standard deviations
%of the random errors in the observations, these errors are in practice inflated
%to account for sparsity of the retrieval database. Our results suggest that this
%may limit the ability of the method to reproduce structure of the retrieval
%manifold leading to a loss of sensitivity.
%
%In addition to the improvement in retrieval accuracy between GPROF and the
%GPROF-NN algorithms, our results showed further improvements in retrieval
%accuracy are obtained when spatial information is included in the retrieval.
%Intuitively, this makes sense since a human observer can easily identify
%characteristic rain structures in PMW imagery. At this point we can not
%make any statements regarding whether or not the patters learned by the
%neural network have physical significance but we plan to address this in
%a follow up study.
%
%\subsection{The importance of ancillary data}
%
%As described above, GPROF relies on several ancillary inputs to improve
%retrieval performance. Since the GPROF-NN algorithms were designed to be
%equivalent to GPROF, they, too, make use of ancillary data. However, the
%improved performance of the neural network based retrievals raises the question
%to what extent the inclusion of ancillary data is necessary. To answer this
%question we have tested two variants (GPROF-NN 0D NA and GPROF-NN 3D NA) of the
%GPROF-NN 0D and GPROF-NN 3D algorithms that don't make use of ancillary data and
%assessed their performance.
%
%\begin{table}[hbpt!]
%  \centering
%  \caption{Accuracy metrics of the surface precipitation retrieval for the
%    retrievals with and without ancillary data.}
%  \begin{tabular}{lrrrr}
%    \toprule
%        {} &      Bias $[\unit{mm\ h^{-1}}]$ &       MAE $[\unit{mm\ h^{-1}}]$ &       RMSE $[\unit{mm\ h^{-1}}]$ &       MAPE $[\unit{\%}]$ \\
%        Algorithm   &           &           &           &            \\
%        \midrule
%        GPROF       &  0.008458 &  0.079320 & 0.489126  &  66.335999 \\
%        GPROF-NN 0D & -0.001904 &  0.057892 & 0.380972  &  58.060650 \\
%        GPROF-NN 0D NA & \textbf{0.000092} &	0.061978 & 	0.379741 & 	60.272976 \\
%        GPROF-NN 3D & -0.001239 &  \textbf{0.046224} &  0.338950 &  \textbf{48.223190} \\
%        GPROF-NN 3D NA & -0.000270 & 0.048000 &	 \textbf{0.328201} &	51.544792  \\
%        \bottomrule
%  \end{tabular}
%
%  \label{tab:surface_precip_metrics_na}
%\end{table}
%
%The results are presented in Tab.~\ref{tab:surface_precip_metrics_na} and in
%Fig.~\ref{fig:results_scatter_na} - \ref{fig:results_surface_types_na} in the
%appendix. Although omitting the ancillary data has a negative impact on the
%retrieval performance, both retrievals still perform better than GPROF.
%Moreover, the GPROF-NN 3D retrieval without ancillary data performs only
%marginally worse than the retrieval with ancillary data and better than both
%variants of the GPROF-NN 0D retrieval.
%
%Although this may not have any direct consequences on the operational
%application of GPROF, it still has interesting consequences for the development
%of future retrievals. Firstly, this obviously means that incorporating spatial
%information into a PMW precipitation retrieval improves the retrieval to a
%larger extent than adding ancillary information from secondary sources without
%adding latency due to data availability constraints. Secondly, this also opens
%up the possibility of distributing the retrieval algorithm instead of the
%retrieval and letting the users run the retrieval themselves. The neural network
%model for the GPROF-NN 3D algorithm has a size of about $70\ \unit{MB}$,
%which makes it roughly as large as the level 1C file required to run the
%retrieval. This use case may be interesting for the development of specialized
%retrievals that have a lower general interest but require large amounts of
%storage space. An example would be a high-resolution hydrometeor profile
%retrieval, which may be of interest for case studies but whose output would be
%too large to archive operationally.
%
%
%\section{Conclusions}
%
%The GPROF-NN algorithms and their performance assessment that were presented in
%this study indicate that the retrieval accuracy of the PMW precipitation retrievals
%of the GPM constellation can be improved significantly by replacing the current
%algorithm with a neural network based implementation. Although the presented
%results should be interpreted as an upper bound on the true retrieval
%performance, our results remain encouraging: Replacing the retrieval method has
%the potential to be a simple and cheap way to improve global precipitation
%measurements.
%
%The next step towards such an upgrade will be to extend the implementation of
%the GPROF-NN 0D and GPROF-NN 3D algorithms to the other sensors of the GPM
%constellation and run them operationally along side GPROF. The three methods
%will then be validated and compared against gauge corrected ground based radar
%measurements in a follow up study. This is to ensure that the good performance of
%the neural network based retrieval does indeed carries over to the operational
%processing.
%
%In addition to potential improvements in retrieval accuracy, we presented two
%novel applications enabled by neural-network based retrievals:
%
%\begin{enumerate}
%\item The inclusion of random samples from the posterior distribution in the
%  retrieval output is proposed as an efficient way of representing retrieval
%  uncertainties that helps to improve the representation of extreme events in
%  the retrieval results.
%\item Our results show that incorporating structural information into the
%  retrieval yields larger performance benefits than incorporating ancillary
%  information. This allows simplifying the retrieval pipeline to the point that
%  only observations are required as input, which opens up the possibility
%  letting the user run the retrieval on his own computer, i.e. to perform the
%  inference 'on the edge'. This may be beneficial for retrievals with very high
%  resolution, i.e. whose storage requirement are much larger than those of the
%  L1C data.
%\end{enumerate}
%
%Building on the approach proposed in \citet{pfreundschuh18}, this work presents
%a way of integrating and reconciling machine learning methods with retrieval
%pipelines built on Bayesian retrieval algorithms. We find that the machine
%learning based retrieval, is easiert to implement, runs faster and performs
%better than the hand-crafted retrieval algorithm. Although it remains to be seen
%to which extent this will improve retrievals with respect to independent
%validation data, it seems reasonable to expect at least incremental improvements
%in global precipitation measurements as well as interesting opportunities for
%the novel retrieval applications.

\section{Appendix}

\begin{table}[hbpt!]
  \centering
  \caption{
    Same as Tab.~\ref{tab:metrics_gmi} but for the MHS sensor.
  }
  \label{tab:metrics_mhs}
\begin{subtable}{\textwidth} 
 \caption{Surface precipitation} 
\begin{tabular}{|c||p{3.5cm}|p{3.5cm}|p{3.5cm}|}
\hline
Metric &
\multicolumn{1}{|c}{GPROF} &
\multicolumn{1}{|c}{GPROF-NN 1D} &
\multicolumn{1}{|c|}{GPROF-NN 3D} \\
\hline\hline
Bias \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0061 \pm 0.0001$ &\hfill $ -0.0027 \pm 0.0001$ &\hfill $ -0.0020 \pm 0.0001$ \\
MAE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0898 \pm 0.0001$ &\hfill $  0.0618 \pm 0.0000$ &\hfill $  0.0502 \pm 0.0001$ \\
RMSE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.2570 \pm 0.0001$ &\hfill $  0.2032 \pm 0.0001$ &\hfill $  0.1289 \pm 0.0001$ \\
MAPE$_{0.1}$ \hfill [$\unit{\%}$] & \hfill $ 81.8390 \pm 0.0782$ &\hfill $ 68.5625 \pm 0.0472$ &\hfill $ 59.1714 \pm 0.0668$ \\
\hline
\end{tabular}
\end{subtable}

\begin{subtable}{\textwidth} 
 \caption{Convective precipitation} 
\begin{tabular}{|c||p{3.5cm}|p{3.5cm}|p{3.5cm}|}
\hline
Metric &
\multicolumn{1}{|c}{GPROF} &
\multicolumn{1}{|c}{GPROF-NN 1D} &
\multicolumn{1}{|c|}{GPROF-NN 3D} \\
\hline\hline
Bias \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0068 \pm 0.0001$ &\hfill $ -0.0045 \pm 0.0001$ &\hfill $  0.0027 \pm 0.0002$ \\
MAE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0366 \pm 0.0001$ &\hfill $  0.0268 \pm 0.0001$ &\hfill $  0.0225 \pm 0.0002$ \\
RMSE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.1760 \pm 0.0002$ &\hfill $  0.1459 \pm 0.0002$ &\hfill $  0.0931 \pm 0.0002$ \\
MAPE$_{0.1}$ \hfill [$\unit{\%}$] & \hfill $ 86.1388 \pm 0.3012$ &\hfill $ 77.8337 \pm 0.1587$ &\hfill $ 93.3863 \pm 0.7875$ \\
\hline
\end{tabular}
\end{subtable}

\begin{subtable}{\textwidth} 
 \caption{Rain water path} 
\begin{tabular}{|c||p{3.5cm}|p{3.5cm}|p{3.5cm}|}
\hline
Metric &
\multicolumn{1}{|c}{GPROF} &
\multicolumn{1}{|c}{GPROF-NN 1D} &
\multicolumn{1}{|c|}{GPROF-NN 3D} \\
\hline\hline
Bias \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0035 \pm 0.0000$ &\hfill $ -0.0000 \pm 0.0000$ &\hfill $  0.0008 \pm 0.0001$ \\
MAE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0231 \pm 0.0001$ &\hfill $  0.0150 \pm 0.0000$ &\hfill $  0.0121 \pm 0.0000$ \\
RMSE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0176 \pm 0.0000$ &\hfill $  0.0129 \pm 0.0000$ &\hfill $  0.0072 \pm 0.0000$ \\
MAPE$_{0.01}$ \hfill [$\unit{\%}$] & \hfill $121.7047 \pm 0.2060$ &\hfill $ 86.6833 \pm 0.1344$ &\hfill $ 79.5081 \pm 0.4045$ \\
\hline
\end{tabular}
\end{subtable}

\begin{subtable}{\textwidth} 
 \caption{Ice water path} 
\begin{tabular}{|c||p{3.5cm}|p{3.5cm}|p{3.5cm}|}
\hline
Metric &
\multicolumn{1}{|c}{GPROF} &
\multicolumn{1}{|c}{GPROF-NN 1D} &
\multicolumn{1}{|c|}{GPROF-NN 3D} \\
\hline\hline
Bias \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0055 \pm 0.0001$ &\hfill $ -0.0006 \pm 0.0000$ &\hfill $  0.0008 \pm 0.0001$ \\
MAE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0274 \pm 0.0000$ &\hfill $  0.0123 \pm 0.0000$ &\hfill $  0.0112 \pm 0.0001$ \\
RMSE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0208 \pm 0.0001$ &\hfill $  0.0105 \pm 0.0001$ &\hfill $  0.0065 \pm 0.0001$ \\
MAPE$_{0.01}$ \hfill [$\unit{\%}$] & \hfill $109.1287 \pm 0.5944$ &\hfill $ 73.1542 \pm 0.0480$ &\hfill $ 73.5670 \pm 0.2495$ \\
\hline
\end{tabular}
\end{subtable}

\begin{subtable}{\textwidth} 
 \caption{Cloud water path} 
\begin{tabular}{|c||p{3.5cm}|p{3.5cm}|p{3.5cm}|}
\hline
Metric &
\multicolumn{1}{|c}{GPROF} &
\multicolumn{1}{|c}{GPROF-NN 1D} &
\multicolumn{1}{|c|}{GPROF-NN 3D} \\
\hline\hline
Bias \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0051 \pm 0.0000$ &\hfill $  0.0015 \pm 0.0000$ &\hfill $  0.0042 \pm 0.0000$ \\
MAE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0293 \pm 0.0000$ &\hfill $  0.0202 \pm 0.0000$ &\hfill $  0.0167 \pm 0.0000$ \\
RMSE \hfill [$\unit{mm\ h^{-1}}$] & \hfill $  0.0032 \pm 0.0000$ &\hfill $  0.0017 \pm 0.0000$ &\hfill $  0.0013 \pm 0.0000$ \\
MAPE$_{0.01}$ \hfill [$\unit{\%}$] & \hfill $ 57.3852 \pm 0.0340$ &\hfill $ 42.5588 \pm 0.0266$ &\hfill $ 37.8377 \pm 0.0820$ \\
\hline
\end{tabular}
\end{subtable}

\end{table}

\clearpage

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=\textwidth]{../figures/results_spatial}
  \caption{
    As \ref{fig:results_spatial} but for MHS.
  }
  \label{fig:results_spatial}
\end{figure}

\begin{figure}[hbpt]
  \centering
  \includegraphics[width=\textwidth]{../plots/timings}
  \caption{
    Single-core processing times of GPROF and the GPROF-NN algorithms for a
    single orbit of observations.
    }
  \label{fig:timings}
\end{figure}

%\begin{figure}[hbpt]
%  \centering
%  \includegraphics[width=\textwidth]{../figures/sensitivity}
%  \caption{
%    Relative gradients with respect to the input observations of the three retrievals.
%    Each panel shows bar plots representing the distribution of the relative gradients
%    for each of the 13 channels and 4 different amounts of surface precipitation (RR).
%    Colored boxes display the interquartile range (IQR), whiskers extent to outermost points that
%    lie within a distance $1.5 \cdot \text{IQR}$ of the limits of the IQR, outliers are
%    not displayed. Note the difference in y-axis range between the results for GPROF and
%    the GPROF-NN retrievals.
%    }
%  \label{fig:sensitivity}
%\end{figure}

%\begin{figure}[hbpt]
%  \centering
%  \includegraphics[width=\textwidth]{../figures/results_scatter_na}
%  \caption{
%    Like Fig.~\ref{fig:results_scatter} but comparing the GPROF-NN algorithms
%    with (GPROF-NN 0D, GPROF-NN 3D) and without ancillary data (GPROF-NN 0D NA,
%    GPROF-NN 3D NA).
%  }
%  \label{fig:results_scatter_na}
%\end{figure}
%
%\begin{figure}[hbpt]
%  \centering
%  \includegraphics[width=\textwidth]{../figures/results_surface_types_na}
%  \caption{
%    Like Fig.~\ref{fig:results_surface_types} but comparing the GPROF-NN algorithms
%    with (GPROF-NN 0D, GPROF-NN 3D) and without ancillary data (GPROF-NN 0D NA,
%    GPROF-NN 3D NA).
%  }
%  \label{fig:results_surface_types_na}
%\end{figure}

\clearpage

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}